Code Club Project: Assessing whether intra and inter-genomic variation hinder utility of ESVs and ASVs

### 2021-01-15 submission/manuscript.tex
Replicate lump/split calculations 100 times

A method to avoid splitting a single genome into multiple units of
inference is to cluster 16S rRNA gene sequences together that are
similar. However, this also increases the risk of lumping together genes
from different species that are similar to each other. Therefore, I
assessed the impact of the threshold used to define clusters of 16S rRNA
genes on the propensity to split a genome apart or to lump species
together. For full length 16S rRNA gene sequences, I found that at a
threshold of \textbf{5.5}\%, 95\% of the species with 7 copies of the
rrn operon would be represented by a single OTU. Similarly, thresholds
of \textbf{2.5}, \textbf{4.0}, and \textbf{3.5}\% were observed for the
V4, V3-V4, and V4-V5 regions, respectively. However, at these
thresholds, multiple species could be represented by the same OTU. At
the highest level of resolution, XX\% of the species shared a 16S rRNA
gene sequence variant with another species. Given the risk of splitting
a genome into multiple OTUs is more biologically problematic than
lumping species together, larger thresholds are advisable.

>>>>

A method to avoid splitting a single genome into multiple units of
inference is to cluster 16S rRNA gene sequences together that are
similar. However, this also increases the risk of lumping together genes
from different species that are similar to each other. Therefore, I
assessed the impact of the threshold used to define clusters of 16S rRNA
genes on the propensity to split a genome apart or to lump species
together. For full length 16S rRNA gene sequences, I found that at a
threshold of \textbf{5.5}\%, 95\% of the species with 7 copies of the
rrn operon would be represented by a single OTU. Similarly, thresholds
of \textbf{2.5}, \textbf{4.0}, and \textbf{3.5}\% were observed for the
V4, V3-V4, and V4-V5 regions, respectively. However, at these
thresholds, multiple species could be represented by the same OTU. At
the highest level of resolution, \textbf{3.6}\% of the species shared a
16S rRNA gene sequence variant with another species when considering
full length sequences and \textbf{14.9}, \textbf{10.2}, and
\textbf{12.0}\% when considering the V4, V3-V4, and V4-V5 regions,
respectively. At the commonly used 3\% threshold, \textbf{25.2}\% of the
species shared an OTU when considering full length sequences and
\textbf{33.0}, \textbf{29.4}, and \textbf{32.2}\% when considering the
V4, V3-V4, and V4-V5 regions, respectively. Given the risk of splitting
a genome into multiple OTUs is more biologically problematic than
lumping species together, larger thresholds are advisable.
